---
title: "Mini Project 3"
author: "Angelica Estrada & Natalia Kreciglowa"
date: "11/17/2017"
output: html_document
---
 

```{r, setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("mdsr")
library(mdsr)
library(lubridate) # for converting columns to `date` types
db <- dbConnect_scidb(dbname = "imdb")
class(db)
library(DBI)
rename <- dplyr::rename
library(dplyr)
```

<!-- ***SQL QUERY*** -->

```{r, include=FALSE}
# Get the Genre of Every Movie:
genre <- db %>%
  dbGetQuery("SELECT movie_id, info_type_id, info, id 
              FROM movie_info
              WHERE info_type_id = 3;")
# save(genre, file = "genre.Rda")
```
 

```{r, include=FALSE}
# Get the Release Date of Every Movie:
release <- db %>%
  dbGetQuery("SELECT movie_id, info_type_id, info, id 
              FROM movie_info
              WHERE info_type_id = 16;")
# save(release, file = "release.Rda")
```


```{r, include=FALSE}
# Get the Title of Every Movie:
title <- db %>%
  dbGetQuery("SELECT id, title 
             FROM title;")
# save(title, file = "title.Rda")
```

<!-- ***DATA CLEANING IN R*** -->

```{r, include=FALSE}
# Rename Id in Title:
title <- title %>%
  rename( movie_id = id )
```


```{r, include=FALSE}
# Left join release date to title:
release <- release %>%
  left_join( title, by = "movie_id")
```

```{r, include=FALSE}
# test duplication deletion:
library(dplyr)
datetitle <- distinct(release, title, .keep_all = TRUE)
#length(unique(duptitle$title))
#length(unique(title$title))
#length(unique(release$title))
#length(unique(datetitle$movie_id))
```

```{r, include=FALSE}
# Count the number of genres per movie_id:
genrecount <- genre %>%
  group_by( movie_id ) %>%
  summarise( genrecount = n() ) 
# length(unique(genrecount$movie_id))
# length(unique(genre$movie_id))
```

```{r, include=FALSE}
# Join Date, Title and Genre Count information:
movieinfo <- datetitle %>%
  inner_join(genrecount, by = "movie_id" )
```

<!-- Create `year` column by extracting year from existing data frame -->

```{r, include=FALSE}
# Split `info` column into `country` and `date` 
movieinfo <- movieinfo %>%
  separate(col = info, 
           into = c("country", "date"), 
           sep = "\\:")

# Formatting of values in `date` column (1 of 3 types):
# 1. day month year
# 2. month year
# 3. year
```
    
```{r, include=FALSE}
# Uses lubridate package

# Create new date column in date format
movieinfo <- movieinfo %>%
  mutate(dates = dmy(date))

# Values in the "day month year" were the only rows that were correctly parsed
# Need to fix the remaining 231521 rows that failed to parse
```

<!-- Resolving year issue for values in "day month year" and "year" format -->

```{r, include=FALSE}
# Convert `date` to numeric and save as a vector called `year`
# as.numeric() will:
#   1. Extract the "year" from `date` values in the "day month year" format and convert the character to numeric
#   2. Convert values in the "year" format to numeric
# Values in the "month year" format will be coded as NA but this will be fixed later

yr <- as.numeric(movieinfo$date)
```
    
```{r, include=FALSE}
### Code below just ensures that `yr` contains only years 
### (skip this chunk but don't delete)

# #count number of `NA`'s in `date`
# sum(is.na(yr)) # 712118 NA's
# 
# #count number of years not including NA
# length(yr)-sum(is.na(yr)) #147324 elements
#  
# # Create vector only containing year and no NA
# yronly <- na.omit(yr) 
# length(yronly) #147324 elements
# 
# # Check that all values in `yronly` are greater than 1800 (to make sure there are no days)
# length(yronly > 1800) #147324 elements (matches!)
```
    

```{r include=FALSE}
# Add `yr` column to `movieinfo`
movieinfo <- movieinfo %>%
  mutate(yr = yr)

# Split `date` column to `day`, `month`, `year` and while keeping `date` column

d <- movieinfo$date

movieinfo <- movieinfo %>%
  separate(col = date,
           into = c("day", "month", "year"),
           sep = "\\ ")

movieinfo <- movieinfo %>%
  mutate(date=d)

# Here we see that some `day` values are listed as a month, and some `month` values are listed as a year. This will be resolved in the code below.
```

```{r, include=FALSE}
# Convert `day` and `year` columns to numeric
# Any `day` values containing a month or string will be coerced to NA
movieinfo <- movieinfo %>%
  mutate(day = as.numeric(day),
         year = as.numeric(year))

# Combine `yr` and `year` columns into `year` column
movieinfo <- movieinfo %>%
  mutate(year = ifelse(is.na(year) == TRUE, yr, year))
```
    
<!-- Resolving year issue for values in "month year" format -->
 
```{r, include=FALSE}
# Split `movieinfo` into 2 dataframes based on (non)missing years

# First split: all movies with year
yr <- movieinfo %>%
  filter(!is.na(year))

# Second split: all movies without year (year = NA)
missing_yr <- movieinfo %>%
  filter(is.na(year))

# Check: Do the number of rows in `yr` and `missing_yr` add up to 859442? YES!
 nrow(yr) + nrow(missing_yr)
```
    
```{r, include=FALSE}
# Within `missing_yr`, we see that `month` column only contains the year and the `year` column has all NA's

# Replace the NA's in `year` column by values in `month` column  
missing_yr <- missing_yr %>%
  mutate(year = month)
```
    
```{r, include=FALSE}
# Combine both `yr` and `missing_yr` dataframes into a single dataframe called `movies_df` 
movies_df <- rbind(yr, missing_yr)

# Double check for any missing years? 0 missing years, we're good to go!
sum(is.na(movies_df$year)) 
```
   
<!-- ***CLEAN AND FINAL DATAFRAME TO USE: `movies`*** -->     
     
```{r, include=FALSE}
# Select relevant columns from `movies` and save as a clean & final dataframe: `movies` 
# Only considering `year`

movies <- movies_df %>%
  select(movie_id, title, year, genrecount, country)

```
    
<!-- ***PLOTS*** -->


```{r, include=FALSE}
# Create data for US plots only:
usamovies <- movies %>%
  filter( country == "USA")
usamovies$year <- as.numeric(as.character(usamovies$year))
```

```{r, include=FALSE}
# Clean data for observation years:
usamoviesdates <- usamovies %>%
  filter( year >= 1937 ) %>%
  filter( year <= 1957 )
```

```{r, include=FALSE}
# Clean Data for a genre % of total movies per year:
genresplit <- usamoviesdates %>%
  group_by( year ) %>%
  summarize( totalmovies = n()) 
```

```{r, include=FALSE}
# Clean Data for each genre count per year:
genresplit1 <- usamoviesdates %>% 
  filter( genrecount == 1 ) %>%
  group_by( year ) %>%
  summarize( totalgenre1 = n()) 
genresplit2 <- usamoviesdates %>% 
  filter( genrecount == 2 ) %>%
  group_by( year ) %>%
  summarize( totalgenre2 = n()) 
genresplit3 <- usamoviesdates %>% 
  filter( genrecount == 3 ) %>%
  group_by( year ) %>%
  summarize( totalgenre3 = n()) 
genresplit4 <- usamoviesdates %>% 
  filter( genrecount == 4 ) %>%
  group_by( year ) %>%
  summarize( totalgenre4 = n()) 
genresplit5 <- usamoviesdates %>% 
  filter( genrecount == 5 ) %>%
  group_by( year ) %>%
  summarize( totalgenre5 = n()) 
genresplit6 <- usamoviesdates %>% 
  filter( genrecount == 6 ) %>%
  group_by( year ) %>%
  summarize( totalgenre6 = n()) 
genresplit7 <- usamoviesdates %>% 
  filter( genrecount == 7 ) %>%
  group_by( year ) %>%
  summarize( totalgenre7 = n()) 
genresplit8 <- usamoviesdates %>% 
  filter( genrecount == 8 ) %>%
  group_by( year ) %>%
  summarize( totalgenre8 = n()) 
genresplit9 <- usamoviesdates %>% 
  filter( genrecount == 9 ) %>%
  group_by( year ) %>%
  summarize( totalgenre9 = n()) 
```

```{r, include=FALSE}
# Join the different genre count data frames:
genresplit <- genresplit %>%
  full_join( genresplit1, by= "year")
genresplit <- genresplit %>%
  full_join( genresplit2, by= "year") 
genresplit <- genresplit %>%
  full_join( genresplit3, by= "year") 
genresplit <- genresplit %>%
  full_join( genresplit4, by= "year") 
genresplit <- genresplit %>%
  full_join( genresplit5, by= "year") 
genresplit <- genresplit %>%
  full_join( genresplit6, by= "year") 
genresplit <- genresplit %>%
  full_join( genresplit7, by= "year") 
genresplit <- genresplit %>%
  full_join( genresplit8, by= "year") 
genresplit <- genresplit %>%
  full_join( genresplit9, by= "year") 
genresplit[is.na(genresplit)] <- 0
```


```{r, include=FALSE}
# Mutate for proportion:
genresplit <- genresplit %>%
  mutate( "Proportion of Movies with 1 Genre" = totalgenre1 / totalmovies) %>%
  mutate( "Proportion of Movies with 2 Genres" = totalgenre2 / totalmovies) %>%
  mutate( "Proportion of Movies with 3 Genres" = totalgenre3 / totalmovies) %>%
  mutate( "Proportion of Movies with 4 Genres" = totalgenre4 / totalmovies) %>%
  mutate( "Proportion of Movies with 5 Genres"= totalgenre5 / totalmovies) %>%
  mutate( "Proportion of Movies with 6 Genres" = totalgenre6 / totalmovies) %>%
  mutate( "Proportion of Movies with 7 Genres"= totalgenre7 / totalmovies) %>%
  mutate( "Proportion of Movies with 8 Genres" = totalgenre8 / totalmovies) %>%
  mutate( "Proportion of Movies with 9 Genres" = totalgenre9 / totalmovies) 
```

## 1947: United States v. Paramount Pictures, Inc.

This case was a landmark United States Antitrust case, in which movie studios were no longer allowed to own their own movie theaters and changed the way a film was produced, filmed and distributed. Studios were no longer allowed to be vertically integrated, with the studio owning every piece of the movie making process: direct contracts with actors or actresses, directors, movie theaters and much more.

This case shook up the entire movie industry, changing how movies were created. Could this have direct impact on the **number of genres** a movie can have? On the IMBD database, movies are classified by each genre they own, and perhaps, following this case, there was more freedom in the way movies could be made, so movies could have many more genres. 

Below, we graphed the years **1937-1957**, to see the proportion of total movies released in a year with based on number of genres each movie had:

```{r, echo=FALSE, message= FALSE}
# Create clean data frame:
library(reshape2)
library(plotly)
genresplitclean <- genresplit %>%
  select(year,  "Proportion of Movies with 1 Genre",  "Proportion of Movies with 2 Genres",  "Proportion of Movies with 3 Genres",  "Proportion of Movies with 4 Genres",  "Proportion of Movies with 5 Genres",  "Proportion of Movies with 6 Genres", "Proportion of Movies with 7 Genres",  "Proportion of Movies with 8 Genres",  "Proportion of Movies with 9 Genres") %>%
  rename( Year = year )
mgenresplitclean <- melt( genresplitclean, id = c("Year"))
mgenresplitclean <- mgenresplitclean %>%
  rename(GenreCount= variable) %>%
  rename(Percent = value)
p <- ggplot(mgenresplitclean , aes( x= Year, Percent, fill = GenreCount)) +
    geom_bar(position = "fill", stat = "identity") +
    scale_fill_brewer(palette="YlGnBu", name="Number of Genres per Movie") +
    theme(panel.border = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_rect(fill = "white"))
p <- ggplotly(p)
p
```


### A Few Observations: 

There are **no dramatic increase** to the number of genres films had post-1947! More notable is the fact that movies with only ONE genre were made at a higher proportion of the total movies released. Perhaps this shows the effects of this Supreme Court Case: studios could no longer "control/buy-out" one genre, by hiring all the horror writers, producers and directors in house. Instead of limiting a single genre to one movie studio who contracted all the genre-specific talent, the movie industry professionals were free to move around, and produce movies in that specific genre accross many different studios. 

<!-- RESOURCES

GITHUB:
https://github.com/a17estrada/sds192-mp3

Splitting a single column into multiple columns
# https://stackoverflow.com/questions/7069076/split-column-at-delimiter-in-data-frame  (uses tidyr package)

Convert to dates
http://blog.yhat.com/static/pdf/R_date_cheat_sheet.pdf
https://rpubs.com/davoodastaraky/lubridate
https://www.statmethods.net/input/dates.html

-->
